### 2015-07-02 narrative/data-reduction.tex
start to reshape the proposal

\subsubsection{Data Reduction Techniques}

{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.

The tradiional approach to transforming results of scientific computation into
knowledge consists of storing the numerical output of a simulation on disk,
followed by a post-processing or analysis stage that might take place over
several ensuing weeks or months. The overarching objective is to gain insight,
or knowledge.  As already mentioned, the anticipated scale of computations
means that this workflow is simply no longer viable.

>>>>

\subsection{Data Refactoring}

### 2015-07-02 narrative/data-reorg.tex
start to reshape the proposal

\subsubsection{Data Reorganization Techniques}
One of the key element to get efficient data access during reading is the re-organization of the
data to ensure that the data that will be read back will involve minimal seeks on the file system 
along with aggregating similar data together to get optimal performance and placing the data
across multiple disks to get high levels of concurrency during reading. We will create APIs to
allow users to group similar data together in a data-model, and annotations (initially done in the
ADIOS layer) to allow relationships of variables to be expressed. We will then investigate both
system (general) data-re-organization techniques based on the utility of data and the relationships
of the data, and application-specific routines (for our initial set of evaluation test cases).  Once
we can generalize the application specific routines to a set of motifs, we can then allow this piece
of code to propagate to the general routines. In all cases we will investigate re-organizing the data
into an integer (most likely 8) bins, where they fall into the most likely accessed data to the least
likely accessed data. These methods will NOT be used for checkpoint restart data, since this
data needs to be accessed exactly how it is originally in memory.  We generally see that these
techniques are highly useful for analysis and  visualization data, where users can inherently place
their intentions of the data (life time of the data for eviction policies, prioritization of data 
importance).

>>>>

%\subsubsection{Data Reorganization Techniques}
One of the key element to get efficient data access during reading is the re-organization of the
data to ensure that the data that will be read back will involve minimal seeks on the file system 
along with aggregating similar data together to get optimal performance and placing the data
across multiple disks to get high levels of concurrency during reading. We will create APIs to
allow users to group similar data together in a data-model, and annotations (initially done in the
ADIOS layer) to allow relationships of variables to be expressed. We will then investigate both
system (general) data-re-organization techniques based on the utility of data and the relationships
of the data, and application-specific routines (for our initial set of evaluation test cases).  Once
we can generalize the application specific routines to a set of motifs, we can then allow this piece
of code to propagate to the general routines. In all cases we will investigate re-organizing the data
into an integer (most likely 8) bins, where they fall into the most likely accessed data to the least
likely accessed data. These methods will NOT be used for checkpoint restart data, since this
data needs to be accessed exactly how it is originally in memory.  We generally see that these
techniques are highly useful for analysis and  visualization data, where users can inherently place
their intentions of the data (life time of the data for eviction policies, prioritization of data 
importance).

### 2015-07-02 narrative/intro.tex
start to reshape the proposal

\section{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
\label{sec:introduction}

@@

Our approach targets the expected characteristics of exascale storage
hardware. The storage layer will be partitioned into multiple heterogenous
tiers with vastly different performance characteristics. The layers
will be further differentiated by the constraints on capacity
and data lifetime within each layer. Tape archival storage will still maintain
data long term, but access to this data will be orders of magnitude slower
than the next layer and naively accessing data from archival storage will
greatly impact productivity.
%
The overarching idea is that data from applications can be characterized by different
levels of importance and different level of importance over time. 
Codes such as the XGC1 simulation from C. S. Chang (PPPL), which is one of largest users of Leadership Class Facilities (over 300M hours at ANL, NERSC, and ORNL) has launched a series of simulations which need to write out 100 PB of data on the Titan system. Due to resource limitations, the simulation will only write about 10 PB, but
when our team considered not only the time it would take to write out 10 PB in 10 days of run, but also where we would store this, which had to be tape since it would take
over half of the file system, we decided that we couldn't write this amount of data. In other words we had to make decisions of which data to save from the original 100 PB.
The first thing that we had to discover is how much space we could have on the parallel file system at any given time, and how much time it would take to archive 1 PB of data,
which turned out to be 10 days, and we could only store 2 PB of data at once. This quickly led us to 1) decide which pieces of data to save from the large amount of data; in 
other words, we had to prioritize which were the most important pieces. We had to further breakdown the large data into two categories, data which we think we will touch after
the simulation is complete ( after the 10 days of wallclock time was finished), and what we would do with this data (the access pattern of the large data). We then had to 
discover application specific data reduction techniques, and encode that into our solution, on top of ADIOS. We then had to implement a crude discovery system to understand
which pieces of data were stored on the parallel file system and which pieces moved over to PPPL, and which pieces were on tape.  This solution worked well for one day
but after the second day more problems occurred, the variability of the time it took to write the data made the I/O prohibitively costly, because even though we had the
majority of the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. We had no method
to monitor this time and then write less data when the time was going to be longer.   After the simulation is complete we also need to visualize this reduced data, and 
an additional problem appeared; just to get a first look at the visualization data seemed like it could take over a month because some of the timesteps were on tape and other
timesteps were on the file system. We had no method to just look at the data quickly and understand which regions of the simulation to focus on.

@@

\subsection{Reading  Perspective}
\label{subsec:reading-perspective}
From the perspective of the users who want to read in the data from their
simulations, and then operate on this data and possibly write data, we see
that in general these users will be running on much less compute resources
than the users who are running the simulation. One of the important aspects
of this class of user is that they want a reduction of latency and they
require a certain amount of quality of service as well. When someone is
doing interactive analysis and expects that the data they read in is about
10s, but it turns into 1,000s, the user typically tends to either wait
another day for doing their analysis or gets frustrated and reads much less
data than they really need to. Furthermore the user tends to read in just a
small sample of the data without much knowledge if they are missing
important pieces of the data.
\subsection{Research Challenges to be address}
\label{subsec:challenges}
Our research into the storage system and middle ware layers must address
many of these challenges including
\begin{enumerate}
\item How do we describe the user intentions at the API layer and have this
  communicated down to the middle-ware and storage layers?
\item How do we allow users to define their user defined compression
  techniques and have this data read back from the storage system layers?
  Where do we execute the decompression techniques 
\item How do we evaluate the tradeoffs at runtime to guide data placement,
  including the movement of data across the network, across the different
  storage tiers? (data movement and quality tradeoff)
\item Can we use different forms of learning techniques as daemons on the
  system to perform data migration from one storage tier to another? For
  example, if we see that a user is looking at one time slice after another
  for a certain object in their dataset which is in the slower tiers of the
  hierarchy, do we automatically propane up the data which has not been
  requested in the hope that this will be requested data? (prediction and
  prefetching, user defined compression on movement)
\item Can we understand the true need of campaign storage and understand the
  possible impact of the campaign storage if we run a hadoop-like
  file-system instead of a lustre/gpfs file system? (additional layers)
\item Can we limit the amount of data duplication? Data space is going to be
  very constrained on the exascale system so we must ensure that minimal
  copies are made of data, and when data is duplicated, they are removed
  through a garbage collection routine. (space management)
\item Can we build a model to give us the time estimates during the reading
  and writing phases, so that rules can be in place from the user
  perspective to make adaptive decisions. For example, if the file system
  says that reading will take 3 months, users can they place in rules to
  then read in a subset of data, and they can understand which data can be
  read quickly and which pieces will take more time. (estimation)
\item Can we understand how to place code in the system which will allow
  data-regeneration to take place. (regeneration)
\item How does the storage system do a better job in managing request from
  all of the users on a LCF than today? We realize that today users who have
  a better middlware system can often lock other users from getting high
  performance when they are running. If we have the concept of currency
  which is eventually used in the same extent as node-hours, then users will
  have to be able to think about how much storage and how much bandwidth
  they can choose. One question that needs to be understood is if there are
  times when the system sees that there are very few storage system
  resources being used, then the lucky users can get the bandwidth cheaper
  than at times of heavy usage. How can we enforce this? (Fairness)
\item How do we manage all of the metdata not only from the principle
  objects, but from the sub-objects? How well will this scale when we have
  users who can potentially create billions of objects from their
  simulation? (scalability and discovery)
\end{enumerate}

>>>>

\section{Introduction} 
\label{sec:introduction}

@@

Our approach targets the expected characteristics of exascale storage
hardware. The storage layer will be partitioned into multiple heterogenous
tiers with vastly different performance characteristics. The layers
will be further differentiated by the constraints on capacity
and data lifetime within each layer. Tape archival storage will still maintain
data long term, but access to this data will be orders of magnitude slower
than the next layer and naively accessing data from archival storage will
greatly impact productivity.
%
{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.  Our approach in this proposal is to enable researchers
to place more knowledge into the SSIO system so that data can be recast from a bunch of bytes
to information which the system can then place and optimize the time to gain knowledge from exascale simulations.
We first motivate this by a problem our team has been working on in the past six months, and then recast this problem into
a series of research questions and challenges which must be addressed in order to optimize the SSIO system for exascale science.

The traditional approach to transforming results of scientific computation into
knowledge consists of storing the numerical output of a simulation on disk,
followed by a post-processing or analysis stage that might take place over
several ensuing weeks or months. The overarching objective is to gain insight,
or knowledge.  As already mentioned, the anticipated scale of computations
means that this workflow is simply no longer viable.
%
The overarching idea is that data from applications can be characterized by different
levels of importance and different level of importance over time. 
Codes such as the XGC1 simulation from C. S. Chang (PPPL), which is one of largest users of Leadership Class Facilities (over 300M hours at ANL, NERSC, and ORNL) has launched a series of simulations which need to write out 100 PB of data on the Titan system. Due to resource limitations, the simulation will only write about 10 PB, but
when our team considered not only the time it would take to write out 10 PB in 10 days of run, but also where we would store this, which had to be tape since it would take
over half of the file system, we decided that we couldn't write this amount of data. In other words we had to make decisions of which data to save from the original 100 PB.
The first thing that we had to discover is how much space we could have on the parallel file system at any given time, and how much time it would take to archive 1 PB of data,
which turned out to be 10 days, and we could only store 2 PB of data at once. This quickly led us to 1) decide which pieces of data to save from the large amount of data; in 
other words, we had to prioritize which were the most important pieces. We had to further breakdown the large data into two categories, data which we think we will touch after
the simulation is complete ( after the 10 days of wallclock time was finished), and what we would do with this data (the access pattern of the large data). We then had to 
discover application specific data reduction techniques, and encode that into our solution, on top of ADIOS. We then had to implement a crude discovery system to understand
which pieces of data were stored on the parallel file system and which pieces moved over to PPPL, and which pieces were on tape.  This solution worked well for one day
but after the second day more problems occurred, the variability of the time it took to write the data made the I/O prohibitively costly, because even though we had the
majority of the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. We had no method
to monitor this time and then write less data when the time was going to be longer.   After the simulation is complete we also need to visualize this reduced data, and 
an additional problem appeared; just to get a first look at the visualization data seemed like it could take over a month because some of the timesteps were on tape and other
timesteps were on the file system. We had no method to just look at the data quickly and understand which regions of the simulation to focus on.

@@

\subsection{Reading  Perspective}
\label{subsec:reading-perspective}
From the perspective of the users who want to read in the data from their
simulations, and then operate on this data and possibly write data, we see
that in general these users will be running on much less compute resources
than the users who are running the simulation. One of the important aspects
of this class of user is that they want a reduction of latency and they
require a certain amount of quality of service as well. When someone is
doing interactive analysis and expects that the data they read in is about
10s, but it turns into 1,000s, the user typically tends to either wait
another day for doing their analysis or gets frustrated and reads much less
data than they really need to. Furthermore the user tends to read in just a
small sample of the data without much knowledge if they are missing
important pieces of the data.

### 2015-07-02 narrative/proposed-work.tex
Adding some sections
adding more sections
Added feiyi's text and graphics
fixed the includes to work with emacs
minor changes.
merge data description, qos app and API
merged naming and discovery
modified:   ../bib/maltzahn.bib 	modified:   infrastructure.tex 	deleted:    policies.tex 	modified:   proposed-work.tex 	modified:   qos-storage.tex 	deleted:    time-estimation.tex
start to reshape the proposal

\section*{Proposed Work}

\begin{itemize}

\item Add plug-in architecture to Sirocco to support selective data compression
\item Add specific tier destinations command to Sirocco (assuming that the caching mechansism cannot achieve the desired effects).
\item Develop profiling system to determine how a data set is used during preparation runs prior to a capability run to determine how to optimize data placement for capability run analytics.
\item

\end{itemize}

>>>>

\section{Proposed Research and Methods}
\label{subsec:challenges}
Our research into the storage system and middle ware layers must address
many of these challenges including
\begin{enumerate}
\item How do we describe the user intentions at the API layer and have this
  communicated down to the middle-ware and storage layers?
\item How do we allow users to define their user defined compression
  techniques and have this data read back from the storage system layers?
  Where do we execute the decompression techniques 
\item How do we evaluate the tradeoffs at runtime to guide data placement,
  including the movement of data across the network, across the different
  storage tiers? (data movement and quality tradeoff)
\item Can we use different forms of learning techniques as daemons on the
  system to perform data migration from one storage tier to another? For
  example, if we see that a user is looking at one time slice after another
  for a certain object in their dataset which is in the slower tiers of the
  hierarchy, do we automatically propane up the data which has not been
  requested in the hope that this will be requested data? (prediction and
  prefetching, user defined compression on movement)
\item Can we understand the true need of campaign storage and understand the
  possible impact of the campaign storage if we run a hadoop-like
  file-system instead of a lustre/gpfs file system? (additional layers)
\item Can we limit the amount of data duplication? Data space is going to be
  very constrained on the exascale system so we must ensure that minimal
  copies are made of data, and when data is duplicated, they are removed
  through a garbage collection routine. (space management)
\item Can we build a model to give us the time estimates during the reading
  and writing phases, so that rules can be in place from the user
  perspective to make adaptive decisions. For example, if the file system
  says that reading will take 3 months, users can they place in rules to
  then read in a subset of data, and they can understand which data can be
  read quickly and which pieces will take more time. (estimation)
\item Can we understand how to place code in the system which will allow
  data-regeneration to take place. (regeneration)
\item How does the storage system do a better job in managing request from
  all of the users on a LCF than today? We realize that today users who have
  a better middlware system can often lock other users from getting high
  performance when they are running. If we have the concept of currency
  which is eventually used in the same extent as node-hours, then users will
  have to be able to think about how much storage and how much bandwidth
  they can choose. One question that needs to be understood is if there are
  times when the system sees that there are very few storage system
  resources being used, then the lucky users can get the bandwidth cheaper
  than at times of heavy usage. How can we enforce this? (Fairness)
\item How do we manage all of the metdata not only from the principle
  objects, but from the sub-objects? How well will this scale when we have
  users who can potentially create billions of objects from their
  simulation? (scalability and discovery)

\item Add plug-in architecture to Sirocco to support selective data compression
\item Add specific tier destinations command to Sirocco (assuming that the caching mechansism cannot achieve the desired effects).
\item Develop profiling system to determine how a data set is used during preparation runs prior to a capability run to determine how to optimize data placement for capability run analytics.

\end{enumerate}

\input{narrative/data-description} % includes data-desc, apis, and qos-app.

%\subsection{Application Interface}

%\input{narrative/data-desc}
\input{narrative/data-reduction} % includes data-reorg and the api specs move to the above section -
\input{narrative/data-reorg}
%\input{narrative/qos-app}
\input{narrative/placement}
%\input{narrative/policies}
%\input{narrative/apis}

%\subsection{Storage System}

\input{narrative/migration}
\input{narrative/resource-mgmt}
\input{narrative/qos-storage}
%\input{narrative/discovery}
\input{narrative/eviction-metrics}
\input{narrative/naming-service}
%\input{narrative/time-estimation}
\input{narrative/infrastructure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:
